---
title: "{{TITLE}}"
author: "{{AUTHOR}}"
date: today
format:
  docx:
{{REFERENCE_DOC}}
    number-sections: true
    toc: true
params:
  data: ""
  ai: false
---

```{r setup, include=FALSE}
library(ezrsurvey)
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                      fig.width = 6.5, fig.height = 3.8, dpi = 150)

# Render with your own data:  quarto render <file> -P data:my-survey.csv
# For multi-file exports, replace this with read_folder("data/") %>% ...
d <- if (nzchar(params$data)) {
  readr::read_csv(params$data, show_col_types = FALSE)
} else {
  podracing_survey
}
ctx <- ai_context(d)
```

# Executive summary

Replace this paragraph with the story of the survey in four sentences: the
headline number, what drives it, the one surprise, and what you recommend
doing about it. Write it last -- it is the only section many readers open.

```{r}
#| eval: !expr params$ai
#| output: asis
# AI draft (render with -P ai:true; needs ellmer + a stored key).
# Rewrite before circulating -- it is a draft, not a finding.
cat(ai_summarise(calc_nps(d, nps_value), template = "exec_summary",
                 context = ctx))
```

# Who answered

Describe the achieved sample: who it represents, who is missing, and anything
that skews it. The tables below carry the exact figures; keep the prose for
what the reader should notice in them.

```{r}
calc_percentage(d, demo_gender, sort = "desc") %>% knitr::kable()
```

```{r}
calc_percentage(d, demo_edu) %>% knitr::kable()
```

# Satisfaction and NPS

State the score and its context (previous wave, sector norm). One chart, one
table, one paragraph of interpretation.

```{r}
plot_nps_gauge(calc_nps(d, nps_value)$nps)
```

```{r}
calc_nps(d, nps_value, by = region) %>% knitr::kable()
```

# What drives satisfaction

Read the matrix by quadrant: top-left (important, delivered poorly) is the
to-do list; top-right (important, delivered well) is what to protect.

```{r}
ipm_model(d, nps_value, "ratings_") %>% plot_ipm()
```

```{r}
#| eval: !expr params$ai
#| output: asis
cat(ai_summarise(ipm_model(d, nps_value, "ratings_"),
                 template = "drivers_barriers", context = ctx))
```

# In their own words

A handful of diverse verbatims puts flesh on the numbers -- illustrations,
not measurements.

```{r}
sample_comments_diverse(d, nps_com, show_com, n = 8, seed = 42) %>%
  dplyr::select(source, comment) %>%
  knitr::kable()
```

# Methodology and precision

```{r}
precision_summary(d, demo_gender, dplyr::starts_with("ratings_"))
```

```{r}
#| eval: !expr params$ai
#| output: asis
cat(ai_summarise(diagnose(d, demo_gender, dplyr::starts_with("ratings_")),
                 template = "methodology", context = ctx))
```
