---
title: "{{TITLE}}"
author: "{{AUTHOR}}"
date: today
format:
  pptx:
{{REFERENCE_DOC}}
params:
  data: ""
---

```{r setup, include=FALSE}
library(ezrsurvey)
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                      fig.width = 12, fig.height = 4.9, dpi = 150)

# Render with your own data:  quarto render <file> -P data:my-survey.csv
# For multi-file exports, replace this with read_folder("data/") %>% ...
d <- if (nzchar(params$data)) {
  readr::read_csv(params$data, show_col_types = FALSE)
} else {
  podracing_survey
}
```

## Executive summary

- Replace these with the three things the room must remember
- One number each, with what it means for the business
- Write them last, once the rest of the deck is settled

::: {.notes}
One message per slide. Replace this deck's placeholder titles with takeaway
titles ("X drives satisfaction", not "Results"): the title row is the only
line most of the room reads.
:::

# The headline

## Net Promoter Score

```{r}
plot_nps_gauge(calc_nps(d, nps_value)$nps)
```

## What drives the score

```{r}
ipm_model(d, nps_value, "ratings_") %>% plot_ipm()
```

# Who answered

## Gender

```{r}
calc_percentage(d, demo_gender, sort = "desc") %>% plot_bars()
```

## Education

```{r}
calc_percentage(d, demo_edu, sort = "asc") %>% plot_bars(flip = TRUE)
```

## Why they attend

```{r}
calc_percentage_multi(d, "motivations_", id = respondent_id, sort = "asc") %>%
  plot_bars(label = option, flip = TRUE)
```

# Appendix

## How to read this report

```{r}
#| output: asis
prec <- precision_summary(d, demo_gender, dplyr::starts_with("ratings_"))
cat(paste0("- ", prec$bullets), sep = "\n")
```
