---
title: "{{TITLE}}"
author: "{{AUTHOR}}"
date: today
format:
  html:
    toc: true
    toc-location: left
    theme: cosmo
    embed-resources: true
    fig-format: svg
params:
  data: ""
---

```{r setup, include=FALSE}
library(ezrsurvey)
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                      fig.width = 9, fig.height = 5, dpi = 150)

# Render with your own data:  quarto render <file> -P data:my-survey.csv
# For multi-file exports, replace this with read_folder("data/") %>% ...
d <- if (nzchar(params$data)) {
  readr::read_csv(params$data, show_col_types = FALSE)
} else {
  podracing_survey
}
```

## Executive summary

Replace this paragraph with the story of the survey in four sentences: the
headline number, what drives it, the one surprise, and what you recommend
doing about it. Write it last -- it is the only section many readers open.

## Who answered

Describe the achieved sample here: who it represents, who is missing, and
anything that skews it (a young panel, one dominant region, ...).

```{r}
calc_percentage(d, demo_gender, sort = "desc") %>% plot_bars()
```

```{r}
calc_percentage(d, demo_edu, sort = "asc") %>% plot_bars(flip = TRUE)
```

```{r}
calc_percentage_multi(d, "motivations_", id = respondent_id, sort = "asc") %>%
  plot_bars(label = option, flip = TRUE)
```

## Satisfaction and NPS

State the score, how it compares to whatever benchmark you have (last wave,
sector norm), and where the promoters and detractors sit.

```{r}
plot_nps_gauge(calc_nps(d, nps_value)$nps)
```

```{r}
plot_nps(d, nps_value)
```

## What drives satisfaction

Read the matrix by quadrant: top-left (important, delivered poorly) is the
to-do list; top-right (important, delivered well) is what to protect.

```{r}
ipm_model(d, nps_value, "ratings_") %>% plot_ipm()
```

## In their own words

A handful of diverse verbatims puts flesh on the numbers. These are
illustrations, not measurements -- say so in the caption.

```{r}
sample_comments_diverse(d, nps_com, show_com, n = 8, seed = 42) %>%
  plot_quotes_tree()
```

## Methodology and precision

```{r}
precision_summary(d, demo_gender, dplyr::starts_with("ratings_"))
```
