# R fundamentals {#fundamentals}

Before the ezr shortcuts, a little groundwork. This chapter is the "what even is
R" chapter. If you have coded before, skim it; if not, take it slowly and type
everything out.

## R and RStudio

**R** is the language. **RStudio** is the comfortable workshop you write it in —
a script editor, a console, a pane showing your data, and a help pane. You will
practically always open RStudio and never touch R directly.

The console is a calculator you talk to in R:

```{r}
1 + 1
10 * 5
seq(0, 100, 25)
```

You can also ask it yes/no questions, which come back as `TRUE`/`FALSE`:

```{r}
10 > 5
5 == 10
```

Anything after a `#` is a **comment** — a note to yourself that R ignores. Use
them generously.

## Values and vectors

You store things in **objects** with `<-` (or `=`). A name must start with a
letter:

```{r}
a <- 5
b <- 10
a + b
```

A **vector** is a sequence of values of the same type, built with `c()`:

```{r}
ratings <- c(4, 5, 3, 5, 2)
mean(ratings)
ratings[1]      # the first element, with square brackets
```

A **factor** is how R stores categories with a defined order — useful for Likert
scales:

```{r}
answers <- c("Disagree", "Agree", "Neutral", "Agree")
factor(answers, levels = c("Disagree", "Neutral", "Agree"))
```

## Data frames

A **data frame** is a table: columns of (possibly different) types, all the same
length. This is what a survey looks like in R. The `theme_park` data is a data
frame:

```{r}
class(theme_park)
nrow(theme_park)
names(theme_park)
```

A handful of functions tell you what you are holding:

```{r}
str(theme_park[, 1:7])     # structure: column types and a peek
summary(theme_park$nps)    # a numeric summary of one column
unique(theme_park$region)  # the distinct values of a column
```

Pull out a column with `$`, and slice rows/columns with `[rows, columns]`:

```{r}
theme_park$nps[1:5]              # first five NPS scores
theme_park[1:3, c("region", "nps")]
```

## A first taste of analysis

Base R already does real statistics. Counts and proportions of a category:

```{r}
table(theme_park$visitor_type)
round(prop.table(table(theme_park$visitor_type)) * 100)
```

Averages, spread, and a quick correlation between two numbers:

```{r}
mean(theme_park$spend)
sd(theme_park$spend)
cor(theme_park$age, theme_park$spend)
```

This is genuinely useful — and also a little clunky. Look at that nested
`round(prop.table(table(...)))` just to get percentages. Hold that thought.

## Why ezr

Here is the same "percentage of each visitor type" in the style we will use for
the rest of the book:

```{r}
calc_percentage(theme_park, visitor_type)
```

One line, a tidy table, the percentages already computed. That is the whole
promise of the ezr family: the analysis you do every day, without the
boilerplate. The next chapter sets it up properly.

<div class="turn">
**Your turn.** Warm up on the fundamentals and first ezr moves:

```{r eval=FALSE}
library(ezrlearning)
x <- draw_exercise(chapter = 2)   # percentages, NPS, the pipe
x
check_answer(x, "your answer")
```
</div>

<div class="recap">
**In a nutshell.**

- RStudio is where you write R; the console is a calculator you instruct.
- Objects hold values; `c()` builds vectors; factors store ordered categories.
- A **data frame** is your survey: `$` grabs a column, `[rows, cols]` slices.
- `table()`, `mean()`, `cor()` already do statistics — but the ezr helpers do the
  everyday ones in a single, readable line.
</div>
