# The ezr way {#the-ezr-way}

R's real power comes from **packages** — bundles of ready-made functions. You
install a package once and load it each session.

```{r eval=FALSE}
# install once (from the family repository)
# install.packages(c("ezrsurvey", "ezrmodel", "ezrlearning"))

# load every session
library(ezrsurvey)
library(ezrmodel)
library(ezrlearning)
```

The single most useful idea you will borrow from the wider tidyverse is the
**pipe**, `%>%` (shortcut: Ctrl/Cmd + Shift + M). It takes the thing on the left
and feeds it into the function on the right, so code reads top-to-bottom like a
sentence:

```{r}
theme_park$visitor_type %>%
  table() %>%
  prop.table()
```

Read it as: *take visitor_type, tabulate it, then turn that into proportions.*

## A default dataset

Most of the time you work with one survey for a while. `ezrsurvey` lets you set
it once with `use_dataset()`, and then the helpers know what you mean without you
repeating the data every time:

```{r}
use_dataset(theme_park)
```

From here on, `calc_percentage(visitor_type)` and
`calc_percentage(theme_park, visitor_type)` mean the same thing. We will mostly
write the data out in full so the examples stand alone, but the shortcut is there
when you want it.

## Percentages without the fuss

The everyday question — *what share of respondents is each category?* — is one
call:

```{r}
calc_percentage(theme_park, region)
```

You get the count (`n`) and the rounded percentage (`pct`). Add `by =` to split
by another variable, and `sort =` to order the result:

```{r}
calc_percentage(theme_park, visitor_type, by = region, sort = "desc")
```

Compare that to the base-R version (`prop.table(table(...))`, rounding,
reshaping) and you can see why we bother.

## NPS in one line

Net Promoter Score is everywhere in insights work: the percentage of *promoters*
(9–10) minus the percentage of *detractors* (0–6). Doing it by hand is fiddly;
`calc_nps()` just does it:

```{r}
calc_nps(theme_park, nps)
```

And, like the percentage helper, it splits by a group when you ask:

```{r}
calc_nps(theme_park, nps, by = region)
```

If you ever need the promoter/passive/detractor label for each respondent,
`nps_group()` gives it:

```{r}
table(nps_group(theme_park$nps, labels = TRUE))
```

## Crosstabs

Two categories at once — a classic contingency table — is `crosstab()`:

```{r}
crosstab(theme_park, visitor_type, region)
```

By default it counts; ask for row or column percentages with `cell = "row_pct"`
or `"col_pct"`, or summarise a third numeric variable with `value =` and `fn =`.

<div class="turn">
**Your turn.** Practise the core verbs — `calc_percentage()`, `calc_nps()`,
`crosstab()`:

```{r eval=FALSE}
library(ezrlearning)
q <- quiz(4, chapter = 2, seed = 2)
q                       # four questions
# work through them, then:
grade(q, list("A", "calc_nps(theme_park, nps, by = region)", "B", "A"))
```
</div>

<div class="recap">
**In a nutshell.**

- `library()` loads a package; `%>%` pipes data left-to-right through functions.
- `use_dataset()` sets a working survey so you stop repeating it.
- `calc_percentage()`, `calc_nps()` and `crosstab()` replace whole chains of
  `count()` / `prop.table()` / `pivot_wider()` with one readable line — and they
  all take a `by =` grouping.
</div>
