# Data and graphs {#data-and-graphs}

Real survey data arrives messy. Rating questions come back as words, ages as free
text, "Prefer not to say" sprinkled through. This chapter is about getting data
into shape and then into a chart you can put in a report.

## Cleaning worded ratings

Look at the rating columns in `theme_park`:

```{r}
unique(theme_park$rating_rides)
```

They are words, not numbers — you cannot average "Very good". `recode_likert()`
turns a worded 5-point scale into the numbers 1–5:

```{r}
recode_likert(theme_park$rating_rides)[1:8]
```

If a column has numbers hidden inside text (like `"7 out of 10"`),
`ensure_numeric()` salvages them:

```{r}
ensure_numeric(c("7 out of 10", "rated 9", "8"))
```

And `na_blank()` turns non-answers (blanks, "Prefer not to answer") into proper
missing values so they do not pollute your averages:

```{r}
na_blank(c("Good", "", "Prefer not to answer", "Bad"))
```

Cleaning all six rating columns at once is a tidy one-liner with `mutate()` and
`across()`:

```{r}
parks <- theme_park %>%
  dplyr::mutate(dplyr::across(dplyr::starts_with("rating_"), recode_likert))
parks[, c("rating_rides", "rating_food", "rating_value")] %>% head(4)
```

Now they are numbers, ready to average, correlate or model.

## Ages into generations

Insights work loves a generational cut. `recode_generation()` maps an age (or a
birth year) to its cohort:

```{r}
theme_park$generation <- recode_generation(theme_park$age, input = "age")
calc_percentage(data.frame(generation = theme_park$generation), generation)
```

(There is `recode_age()` too, for fixed age bands.)

## A first chart

Charts in ezr follow the same one-line spirit. Take a percentage table and pipe
it into `plot_bars()`:

```{r}
calc_percentage(theme_park, region, sort = "desc") %>%
  plot_bars(label = region, title = "Where our visitors come from")
```

For NPS there is a purpose-built gauge-style chart:

```{r}
plot_nps(theme_park, nps, title = "Overall NPS")
```

## Importance vs performance

Here is the chart that earns its keep in a satisfaction study. Which aspects of
the visit matter most for recommending the park (**importance**), and how well do
we score on them (**performance**)? `ipm_model()` works it out — and because the
ratings are still worded, we let it recode them for us:

```{r}
ipm <- ipm_model(theme_park, nps, "rating_")
ipm
```

Each rating gets an *importance* (how strongly it correlates with NPS) and a
*performance* (its average), plus a BAD / OK / GOOD band. Plot it:

```{r, fig.height=4.5}
plot_ipm(ipm, title = "What drives recommendation at the park")
```

Items that are important **and** weak are your priorities; important and strong
are your strengths to protect. That single chart often *is* the headline of a
satisfaction report.

<div class="turn">
**Your turn.** Practise cleaning, generations and the importance model:

```{r eval=FALSE}
library(ezrlearning)
x <- draw_exercise(chapter = 3, type = "code", seed = 4)
x
check_answer(x, "recode_likert(theme_park$rating_value)")
```
</div>

<div class="recap">
**In a nutshell.**

- `recode_likert()` turns worded scales into 1–5; `ensure_numeric()` rescues
  numbers from text; `na_blank()` removes non-answers.
- `across()` inside `mutate()` cleans many columns at once.
- `recode_generation()` cuts age into cohorts.
- `plot_bars()`, `plot_nps()` and `plot_ipm()` turn tidy tables into
  report-ready charts; `ipm_model()` + `plot_ipm()` is the importance–performance
  workhorse.
</div>
