# Finding patterns {#finding-patterns}

So far we have described the data. Now we ask *why*: what explains a respondent's
recommendation? This is where `ezrmodel` comes in, and where the family's
**target-variable** mentality pays off — point a helper at an outcome and let it
tell you what relates to it.

First, a numeric table to model on. We recode the worded ratings to 1–5 and keep
the outcome and a couple of numeric extras:

```{r}
parks <- theme_park %>%
  dplyr::mutate(dplyr::across(dplyr::starts_with("rating_"), recode_likert)) %>%
  dplyr::rename_with(~ sub("^rating_", "", .x), dplyr::starts_with("rating_")) %>%
  dplyr::select(nps, rides, food, staff, cleanliness, value, queue)
head(parks)
```

## What drives the target?

`drivers()` is the signature move. Point it at a target and it ranks every
predictor by **several methods at once** — plain correlation, a regression
coefficient, relative weights, random forest, a common factor — then combines
them into a consensus. Agreement across methods is a robust signal.

```{r}
d <- drivers(parks, nps)
d
```

The consensus tells you which ratings matter most for recommendation. A picture:

```{r}
plot(d)
```

## Correlations, target-focused

If you just want the correlations with the outcome, `correlations()` gives a
tidy, ranked table:

```{r}
correlations(parks, nps)
```

## A regression you can read

`model_lm()` fits an ordinary regression and hands back a tidy result — tidy
coefficients, fit statistics, the model itself — instead of the wall of text
`summary()` prints:

```{r}
fit <- model_lm(parks, nps ~ rides + food + staff + value)
fit
```

Want the model to choose its own predictors? `model_select()` does stepwise or
penalised selection:

```{r}
step <- model_select(parks, nps ~ rides + food + staff + cleanliness + value + queue,
                     method = "stepwise")
step
```

And `compare_models()` lines candidates up so you can pick between them at a
glance:

```{r}
compare_models(simple = fit, stepwise = step)
```

<div class="turn">
**Your turn.** Practise drivers, correlations and model selection:

```{r eval=FALSE}
library(ezrlearning)
q <- quiz(3, chapter = 4, seed = 6)
q
grade(q, list("A", "correlations(parks, nps)", "A"))
```
</div>

<div class="recap">
**In a nutshell.**

- Recode worded ratings to numbers, then point helpers at the **target**.
- `drivers()` ranks predictors by a consensus of several methods — the robust
  answer to "what matters most".
- `correlations()` gives a target-focused ranking; `model_lm()` fits a readable
  regression; `model_select()` picks predictors for you; `compare_models()`
  chooses between fits.
</div>
