# Segments and structure {#segments-and-structure}

Sometimes you are not predicting an outcome — you want to find the *groups* and
the *hidden dimensions* in your data. Segmentation and structure are what this
chapter is about.

We recode the worded ratings to numbers but keep the rest of the survey, so we
can read the segments against demographics later. The rating columns are renamed
(`rating_rides` becomes `rides`):

```{r}
parks <- theme_park %>%
  dplyr::mutate(dplyr::across(dplyr::starts_with("rating_"), recode_likert)) %>%
  dplyr::rename_with(~ sub("^rating_", "", .x), dplyr::starts_with("rating_"))
rating_items <- c("rides", "food", "staff", "cleanliness", "value", "queue")
```

## Clustering with diagnostics

`cluster()` groups respondents by how similarly they rated the park (we point it
at the rating columns with `vars =`). Crucially it returns the *whole* picture in
one object: the assignments, the group sizes, the profiles, **and** the quality
diagnostics. Leave `k` out and it picks the number of clusters for you (by
silhouette width):

```{r}
cl <- cluster(parks, k = 3, vars = c(rides, food, staff, cleanliness, value, queue))
cl
```

The print-out shows how many respondents fell into each segment and how
well-separated they are. A picture places the segments in two dimensions:

```{r}
plot(cl)
```

To read a segment, look at its profile — the average rating per group — which the
object already carries; `cluster_profile()` composes the clusters against another
variable in the data, e.g. visitor type:

```{r}
cluster_profile(cl, visitor_type)
```

## Reducing dimensions

Six rating columns may really be two or three underlying ideas. `reduce_dims()`
runs a principal component analysis and returns the scores, the loadings and the
variance explained:

```{r}
r <- reduce_dims(parks, vars = c(rides, food, staff, cleanliness, value, queue))
r
```

```{r}
plot(r)
```

The scree plot shows how few components capture most of the variation — useful
both to simplify and to sanity-check that your rating battery isn't just noise.

## Do the items hang together?

Before you average a set of items into a scale, check they belong together.
`reliability()` reports Cronbach's alpha:

```{r}
reliability(parks, items = c(rides, food, staff, cleanliness, value, queue))
```

And `factors()` finds the latent dimensions behind the items — which questions
group into the same underlying theme:

```{r}
f <- factors(parks, vars = c(rides, food, staff, cleanliness, value, queue), n = 2)
f
```

```{r}
plot(f)
```

<div class="turn">
**Your turn.** Practise clustering and reliability:

```{r eval=FALSE}
library(ezrlearning)
x <- draw_exercise(chapter = 5, seed = 8)
x
reveal(x)
```
</div>

<div class="recap">
**In a nutshell.**

- `cluster()` segments respondents and returns sizes, profiles and quality
  diagnostics together; it can choose `k` for you.
- `cluster_profile()` reads a segmentation against another variable.
- `reduce_dims()` (PCA) finds the few dimensions behind many columns.
- `reliability()` checks a scale holds together; `factors()` uncovers the latent
  themes in a battery of items.
</div>
