# Text {#text}

Open-ended comments are where the *why* often hides. `ezrmodel` brings the same
one-line spirit to text: tokenise, weigh, find themes, summarise.

Our survey has a free-text `comment` column:

```{r}
theme_park$comment[1:3]
```

## Tokenising

`tokenize_text()` turns the comments into one tidy row per word, lower-cased and
cleaned, with stop-words removed:

```{r}
tk <- tokenize_text(theme_park, comment, respondent_id)
tk
```

A quick count of the most frequent terms:

```{r}
plot(tk)
```

## Distinctive words by group

Raw counts are dominated by words everyone uses. The more interesting question is
which words are *distinctive* to a group. `term_freq()` with a `by` grouping
weights terms by tf-idf — here, the words that set promoters apart from
detractors:

```{r}
tp <- theme_park %>%
  dplyr::mutate(group = nps_group(nps, labels = TRUE))
term_freq(tp, comment, by = group, top = 6)
```

## Finding themes

`topics()` fits a topic model and returns the defining terms of each theme plus
each comment's topic mix — a fast way to see what people talk about without
reading every line:

```{r}
tm <- topics(theme_park, comment, respondent_id, k = 3)
tm
```

```{r, fig.height=4.5}
plot(tm)
```

## An extractive summary

Need a quick, faithful summary in respondents' *own* words? `summarise_text()`
picks the most central sentences (no API key, nothing invented):

```{r}
summarise_text(theme_park, comment, n = 3)
```

<div class="turn">
**Your turn.** Practise the text verbs:

```{r eval=FALSE}
library(ezrlearning)
x <- draw_exercise(chapter = 6, seed = 9)
x
check_answer(x, "term_freq(theme_park, comment)")
```
</div>

<div class="recap">
**In a nutshell.**

- `tokenize_text()` makes a tidy word table; `term_freq(..., by =)` surfaces each
  group's *distinctive* words via tf-idf.
- `topics()` discovers themes and each comment's mix of them.
- `summarise_text()` extracts the most representative sentences verbatim.
</div>
