# Automation and reporting {#automation-and-reporting}

The real payoff of R is doing a thing **once** and then re-running it forever.
This chapter is about writing your own small functions, saving outputs, checking
how trustworthy your numbers are, and drafting narrative.

## Writing your own function

When you find yourself repeating an analysis, wrap it in a function. A function
takes inputs, does something, and returns a result:

```{r}
nps_by <- function(data, group) {
  calc_nps(data, nps, by = {{ group }})
}

nps_by(theme_park, region)
nps_by(theme_park, visitor_type)
```

The `{{ group }}` is how you pass an unquoted column name through to an ezr
helper. Now your house-standard breakdown is one named call you can reuse and
share.

## Saving outputs

The `save_*()` helpers write whatever you have to disk, picking the format from
the file extension:

```{r}
tab <- calc_percentage(theme_park, region)
save_data(tab, file.path(tempdir(), "region.csv"))

p <- plot_nps(theme_park, nps)
save_plot(p, file.path(tempdir(), "nps.png"))
```

To drop several tables into one Excel workbook — a tab each — use `export_xlsx()`:

```{r}
export_xlsx(
  regions = calc_percentage(theme_park, region),
  nps     = calc_nps(theme_park, nps, by = region),
  path    = file.path(tempdir(), "report.xlsx")
)
```

## How precise are these numbers?

Before you put a number in front of a stakeholder, it helps to know how solid it
is. `precision_summary()` gives a plain-language read on sampling precision:

```{r}
precision_summary(theme_park, nps)
```

This is the difference between "NPS is 19" and "NPS is 19, and with this sample
that is a precise estimate" — the second is what earns trust.

## Drafting the narrative

Finally, the family can help you *write*. `ai_summarise()` sends a summary table
(never the raw rows) to a language model and returns a tight paragraph or
bullets. It needs an API key, stored once and kept out of your scripts:

```{r eval=FALSE}
set_llm_key("openai")                       # stored in your OS keyring

ipm_model(theme_park, nps, "rating_") %>%
  ai_summarise(template = "drivers_barriers", provider = "openai")
```

Treat the draft as a *first* draft — you still own the numbers and the judgement.

<div class="turn">
**Your turn.** Practise the reporting verbs:

```{r eval=FALSE}
library(ezrlearning)
x <- draw_exercise(chapter = 7, seed = 10)
x
check_answer(x, "A")
```
</div>

<div class="recap">
**In a nutshell.**

- Wrap repeated analysis in a function; pass columns through with `{{ }}`.
- `save_data()` / `save_plot()` write outputs by file extension; `export_xlsx()`
  puts many tables on their own tabs.
- `precision_summary()` tells you how trustworthy an estimate is.
- `ai_summarise()` drafts the narrative from a summary table (with your key kept
  safe) — a starting point, not the final word.
</div>
