# Practice {#practice}

Reading is not the same as remembering. This last chapter is about the practice
engine built into `ezrlearning` — a way to drill what you have learned, generate
revision sets, and (for tutors) hand out reproducible worksheets.

## Draw an exercise

`draw_exercise()` hands you one question. With a `seed`, it is reproducible — the
same seed always gives the same question, so a class all gets the identical task:

```{r}
x <- draw_exercise(topic = "nps", seed = 1)
x
```

Answer it with `check_answer()` — a letter for multiple choice, or your **actual
ezr code** for a code task (it runs your code and compares the result):

```{r}
check_answer(x, "B")
```

Stuck? `hint(x)` nudges you; `reveal(x)` shows the worked solution:

```{r}
reveal(x)
```

A code task is graded by running what you write:

```{r}
y <- draw_exercise(topic = "nps", type = "code", seed = 2)
y
check_answer(y, "calc_nps(theme_park, nps, by = region)")
```

## Build a quiz

`quiz()` assembles several questions; `grade()` scores them all at once:

```{r}
q <- quiz(4, chapter = 2, seed = 7)
q
```

```{r}
grade(q, list(
  q$exercises[[1]]$answer,            # pretend we answered correctly
  q$exercises[[2]]$answer,
  q$exercises[[3]]$answer,
  q$exercises[[4]]$answer
))
```

## Hand out a worksheet

`export_worksheet()` writes a self-contained R Markdown worksheet that *rebuilds
the identical quiz from its seed* — so every question, with its exact data, is
live inside the document. Add `answer_key = TRUE` for the tutor's copy:

```{r}
path <- export_worksheet(q, file.path(tempdir(), "worksheet.Rmd"),
                         answer_key = TRUE)
basename(path)
```

## Fresh exercises from an LLM

When you want endless variety, `generate_exercise()` drafts new questions with a
language model, pinned to the real ezr functions and the `theme_park` dataset:

```{r eval=FALSE}
set_llm_key("openai")
generate_exercise("NPS by region", type = "mcq")
generate_quiz(5, topic = "clustering")
```

Generated multiple-choice questions are auto-graded; generated code tasks are
self-graded (use `reveal()`), and you should always skim AI output before handing
it to learners.

## What to practise where

Every exercise is tagged with the chapter it belongs to, so you can drill exactly
what you just read. The map:

```{r}
book_chapters()
```

So after Chapter 4 you might run `quiz(5, chapter = 4)`; to revise everything,
`quiz(10)`. Keep a seed if you want to come back to the same set.

<div class="recap">
**In a nutshell.**

- `draw_exercise()` gives one reproducible question; `check_answer()` grades a
  letter or your real ezr code; `hint()` and `reveal()` help.
- `quiz()` + `grade()` handle a whole set; `export_worksheet()` makes a
  self-rebuilding worksheet (with an optional answer key).
- `generate_exercise()` / `generate_quiz()` add AI-authored variety.
- `book_chapters()` connects every topic back to the reading — so practice and
  book stay in step.
</div>

## Where to go next {-}

You now have the whole loop: read a chapter, practise it, and re-run your
analysis whenever the data refreshes. From here, explore each package's own help
and reference site, and keep the family's motto in mind — the analysis you do
every day should take one readable line. Happy analysing.
