---
title: "Learn the ezr family"
subtitle: "Everyday survey analysis and modelling in R, the easy way"
author: "Andreas Schellewald"
date: "`r format(Sys.Date(), '%B %Y')`"
site: bookdown::bookdown_site
documentclass: book
github-repo: aschellewald/ezr
description: "A hands-on introduction to R for research and insights work, taught with the ezr family of packages."
---

# Welcome {-}

```{r setup, include=FALSE}
source("_common.R")
```

This is a friendly, hands-on introduction to using **R** for everyday research
and insights work — the kind of survey analysis, charting and modelling that
fills a research manager's week. It grew out of a series of in-house workshops,
and it keeps that workshop spirit: slow, plain-spoken, lots of small examples,
and no assumption that you have coded before.

What makes this book a little different is the toolkit. Instead of wrestling with
long chains of base-R and tidyverse code, we lean on the **`ezr` family** of
packages, which turn the things you do over and over into single, readable lines:

- [**`ezrsurvey`**](https://github.com/aschellewald/ezr) — load survey exports,
  compute percentages and NPS, clean Likert scales, recode ages into
  generations, build importance–performance charts, and save tidy outputs.
- [**`ezrmodel`**](https://github.com/aschellewald/ezr) — find what drives a
  target variable, cluster customers, reduce dimensions, fit and compare
  regressions, and mine open text.
- [**`ezrlearning`**](https://github.com/aschellewald/ezr) — *this* package: the
  book you are reading plus a practice engine that hands you reproducible
  exercises and grades your answers.

## Who this is for {-}

You are a research, insights or analytics person who works with survey and
customer data, you may live in Excel or a BI tool today, and you would like R to
take the repetitive grind out of your reporting. You do **not** need a maths or
programming background.

## How to read it {-}

Work through it in order — each chapter builds on the last. Type the code out
yourself rather than copying; the muscle memory matters. Every chapter ends with
two things:

- a **Your turn** box pointing you at live practice exercises, and
- an **In a nutshell** recap of the key moves.

The practice exercises come from `ezrlearning` itself. Whenever you see a box
like this:

<div class="turn">
**Your turn.** Try a couple of warm-ups:

```{r eval=FALSE}
library(ezrlearning)
x <- draw_exercise(chapter = 1)   # a random exercise for this chapter
x                                 # read the question
check_answer(x, "B")             # check an answer
reveal(x)                         # see the worked solution
```
</div>

...you can run it straight away. Same `seed`, same question — so a tutor can hand
out `draw_exercise(seed = 42)` and everyone gets the identical task.

## The dataset {-}

Almost every example uses one fictional dataset that ships with the package: a
**theme-park satisfaction survey**, `theme_park`.

```{r}
library(ezrlearning)
head(theme_park)
```

It has the texture of a real survey — a 0–10 recommendation score (`nps`), some
worded rating questions that need cleaning, demographics, spend, and a free-text
comment. We will get to know it well.

Let's begin.
