---
title: "Getting started with ezrintelligence"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with ezrintelligence}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(ezrintelligence)
```

`ezrintelligence` writes the findings that go next to a table. You do the
analysis in whatever package you already use, then hand the finished summary
table to a language model together with a prompt template and the study's
context. Nothing here re-analyses your data: the table you pass is the table
the model sees.

## Store a key once

Keys live in the operating system credential store, never in your scripts.

```r
set_llm_key("openai")          # prompts, without echoing
has_llm_key("openai")
```

A conventional environment variable (`OPENAI_API_KEY`, `ANTHROPIC_API_KEY`,
...) is used as a fallback, and `"ollama"` needs no key at all: point it at a
locally served open-weight model when the tables must not leave the machine.

## Summarise a table

```r
answers <- data.frame(
  answer = c("Very satisfied", "Satisfied", "Neither", "Dissatisfied"),
  pct = c(31, 42, 18, 9)
)

ai_summarise(answers, template = "exec_summary")
```

`template` picks how the answer is written. The registry ships nine:

```{r}
list_prompts()
```

Each one carries a system prompt, a default instruction and an output
contract, all written to the same ground rules: quote only figures present in
the data, treat differences smaller than the margin of error as noise, name
the base. Add your own with `register_prompt()`.

## Give the model the study, not just the table

A percentage without a sample size behind it invites the model to narrate
noise. `ai_context()` bundles the facts an analyst would state alongside the
figures, and derives a margin-of-error note from the sample size:

```{r}
ctx <- ai_context(
  n = 1000,
  question = "How satisfied are you overall?",
  base = "All respondents",
  fieldwork = "12-19 May 2026, online panel"
)
ctx
```

Pass it to every call for the same study:

```r
ai_summarise(answers, template = "key_findings", context = ctx)
```

## Draft a whole report, or a slide

`ai_report_sections()` runs over a named list, one summary per section:

```r
ai_report_sections(
  list(
    satisfaction = answers,
    nps = list(data = nps_table, template = "exec_summary")
  ),
  context = ctx
)
```

`ai_slide_text()` returns the two pieces of text a chart slide needs, a
headline title and three or four takeaway bullets, as plain strings you place
in whatever deck builder you use:

```r
chat <- ai_chat()
text <- ai_slide_text(answers, title = "Satisfaction", chat = chat,
                      context = ctx)
text$title
text$bullets
```

Reusing one `chat` across slides reuses the connection. A failed call messages
and returns `NULL` in that slot, so a network hiccup leaves you with a plain
slide rather than a broken deck.

## Set defaults once

```{r}
ezrintelligence_options()
```

Set them for the session with `ezrintelligence_options(provider =
"anthropic")`, or persist them in a YAML profile with
`use_ezrintelligence_profile()`. Keys are never written to a profile.
