---
title: "Getting started with ezrsurvey"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with ezrsurvey}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 4
)
```

`ezrsurvey` turns the things consumer-survey analysts do over and over into
single-line helpers. This vignette walks through a typical analysis of the
bundled `consumer_survey` dataset: from raw answers to percentages, models,
diagnostics and a chart you could drop on a slide.

```{r setup}
library(ezrsurvey)
```

## The data

`consumer_survey` is 1,000 simulated respondents with demographics, a 0--10 NPS
question, worded feature ratings, multi-select motivations, brand questions and
open-text comments.

```{r}
consumer_survey[, c("respondent_id", "demo_gender", "demo_age", "nps_value")]
```

## Percentages without the `count`/`mutate`/`pivot_wider` dance

`calc_percentage()` is the workhorse: counts, percentages and ordering in one
call.

```{r}
calc_percentage(consumer_survey, demo_gender, sort = "desc")
```

Group with `by`, and pivot to a wide cross-tab with `wide = TRUE`:

```{r}
calc_percentage(consumer_survey, satis_return, by = region, wide = TRUE)
```

Check-all-that-apply questions live in a block of columns sharing a prefix;
`calc_percentage_multi()` handles the per-respondent denominator for you:

```{r}
calc_percentage_multi(consumer_survey, "motivations_",
                      id = respondent_id, sort = "desc")
```

## Reusable level orders

Rather than retyping `levels = c(...)`, register an order once and link it to a
variable; `calc_percentage()` then applies it automatically.

```{r}
register_order(
  "education",
  levels = c("Less than high school", "High school or equivalent",
             "Some college but no degree", "Associate degree",
             "Bachelor degree", "Masters degree or higher"),
  vars = "demo_edu"
)

calc_percentage(consumer_survey, demo_edu)
```

## Net Promoter Score

```{r}
calc_nps(consumer_survey, nps_value)
```

```{r}
plot_nps(consumer_survey, nps_value)
```

## Importance / performance

`ipm_model()` combines driver importance (relative weights against the NPS
score) with feature performance; `plot_ipm()` draws the matrix.

```{r}
model <- ipm_model(consumer_survey, nps_value, "ratings_")
model
```

## Comparing two events

For event surveys you often compare this wave with the last. `compare_values()`
computes the differences and `plot_diff()` draws the change chart.

```{r}
previous <- transform(model, performance = performance - c(.3, -.1, .2, 0, .1))
compare_values(model, previous) |>
  plot_diff()
```

## How precise is it?

`diagnose()` reports the standard error, relative standard error and margin of
error for any column -- numeric (point) or categorical (proportion).

```{r}
diagnose(consumer_survey, demo_gender, nps_value)
```

For a one-shot, plain-language read on the whole study, `precision_summary()`
picks the point-scale and proportion variables for you and writes the bullet
points you would otherwise hand-craft for a report appendix:

```{r}
precision_summary(consumer_survey)
```

### Why relative standard error?

A standard error on its own is hard to judge -- is `±0.05` good? It depends on
the scale. The **relative standard error** (RSE) divides the standard error by
the estimate, so precision becomes comparable across a 1--5 rating and a
percentage alike. Rating estimates by their RSE is the approach national
statistical agencies use to decide whether a number is solid enough to publish;
the Australian Bureau of Statistics, for example, flags estimates by their RSE
before release. ezrsurvey uses a five-band scheme tuned for consumer surveys:

| RSE | Rating |
|---|---|
| under 5% | high precision |
| under 10% | precise |
| under 15% | satisfactory |
| under 25% | use with caution |
| 25% or more | likely reliability issues |

For proportions, the margin of error is widest at a 50/50 split
(`sqrt(0.5 * 0.5 / n)`), so a single worst-case `±` percentage-point figure at
your achieved sample size bounds every percentage in the study -- which is why
`precision_summary()` reports it.

## Saving and exporting

Every table and plot has a one-line, pipe-friendly save:

```{r, eval = FALSE}
calc_percentage(consumer_survey, demo_gender) |> save_data("gender.xlsx")
plot_nps(consumer_survey, nps_value) |> save_plot("nps.png")

# several tables, one tab each
export_xlsx(
  gender = calc_percentage(consumer_survey, demo_gender),
  job    = calc_percentage(consumer_survey, demo_job),
  path   = "summary.xlsx"
)
```

## Where to next

* Set persistent defaults with `ezrsurvey_options()` and a profile
  (`use_ezrsurvey_profile()`).
* Draft narrative summaries with `ai_summarise()` (needs an API key via
  `set_llm_key()`).
* Build a deck with `report_deck()` or scaffold a Quarto report with
  `scaffold_report()`.

```{r, include = FALSE}
remove_order("education")
```
