---
title: "Pod-Racing Fan Survey"
subtitle: "Annual fan study: findings, drivers and recommendations"
author: "Insights team"
date: today
format:
  html:
    toc: true
    toc-location: left
    theme: cosmo
    embed-resources: true
    fig-format: svg
params:
  ai: false
---

```{r setup, include=FALSE}
library(ezrsurvey)
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      fig.width = 9, fig.height = 5, dpi = 150)

use_dataset(podracing_survey)

ctx <- ai_context(
  podracing_survey,
  question = "Annual fan survey after the pod-racing season",
  fieldwork = "Example data bundled with ezrsurvey"
)

# Every figure quoted in the prose is computed here and interpolated inline, so
# the narrative cannot drift away from the charts when the data changes.
n_total <- nrow(podracing_survey)
nps <- calc_nps(nps_value)$nps
grp <- nps_group(podracing_survey$nps_value)
share_promoter <- round(mean(grp == 1, na.rm = TRUE) * 100)
share_detractor <- round(mean(grp == -1, na.rm = TRUE) * 100)

ipm <- ipm_model(nps_value, "ratings_")
threshold <- bands_rating_3()$to[1]
priority <- ipm[ipm$performance < threshold, ]
priority <- priority[order(-priority$importance), ]
strongest <- ipm[order(-ipm$performance), ][1, ]
top_driver <- ipm[order(-ipm$importance), ][1, ]

regions <- calc_nps(nps_value, by = region)
regions <- regions[order(-regions$nps), ]
best_region <- regions[1, ]
worst_region <- regions[nrow(regions), ]

motiv <- calc_percentage_multi("motivations_", id = respondent_id,
                               sort = "desc")
demo <- calc_percentage(demo_gender, sort = "desc")
mean_age <- calc_summary(demo_age)$mean

ret <- calc_percentage(satis_return)
share_return <- round(sum(ret$pct[ret$satis_return %in%
                                    c("Likely", "Very likely")]))

fmt <- function(x) format(x, big.mark = ",")
```

A study of `r fmt(n_total)` pod-racing fans, covering how they rate the
experience, what drives their willingness to recommend it, and where the
opportunities sit. Code is shown alongside the output so the analysis can be
audited and re-run; for a blank scaffold to start your own report from, use
`scaffold_report()`.

## Executive summary

The fan base recommends pod racing, but narrowly. The Net Promoter Score is
**`r sprintf("%+d", nps)`** on a base of `r fmt(n_total)`: `r share_promoter`%
promoters against `r share_detractor`% detractors, with `r share_return`%
saying they are likely to return next season. This is not a lukewarm crowd so
much as a divided one, and the division is explainable.

Enthusiasm rests on the spectacle. `r round(motiv$pct[1])`% attend for the
`r motiv$option[1]`, and `r strongest$feature` is the best-delivered part of
the experience at `r sprintf("%.2f", strongest$performance)` out of 5. That is
the asset to protect.

What holds the score down is delivery on the basics.
`r nrow(priority)` of the `r nrow(ipm)` rated features fall below the
acceptable threshold of `r threshold`.0:
`r paste(sprintf("%s (%.2f)", priority$feature, priority$performance), collapse = ", ")`.
The most consequential of these is **`r priority$feature[1]`**, which carries
`r round(priority$importance[1])`% of explained importance in the
recommendation model while being rated only
`r sprintf("%.2f", priority$performance[1])`. It is the highest-return fix
available.

The weakness is also regional rather than universal:
`r best_region$region` returns `r sprintf("%+d", best_region$nps)` against
`r worst_region$region` at `r sprintf("%+d", worst_region$nps)`, a spread of
`r best_region$nps - worst_region$nps` points that no single global programme
will close.

```{r}
#| eval: !expr params$ai
#| output: asis
#| echo: false
cat(ai_summarise(calc_nps(nps_value), template = "exec_summary", context = ctx))
```

## The headline: willingness to recommend

`r sprintf("%+d", nps)` places the fan base in positive territory, but well
short of the enthusiasm the sport's own marketing assumes.

```{r}
#| fig-height: 2.5
plot_nps_gauge(nps)
```

The distribution matters more than the single number. A score of
`r sprintf("%+d", nps)` could describe a uniformly tepid audience or a split
one; here it is the latter, with substantial mass at both ends.

```{r}
plot_nps(nps_value)
```

Stated return intent corroborates it: `r share_return`% likely or very likely
to come back, with a sizeable undecided middle that is winnable.

```{r}
calc_percentage(satis_return, sort = "desc") %>% plot_bars()
```

### Where the score differs

```{r}
regions
```

`r best_region$region` leads on advocacy at `r sprintf("%+d", best_region$nps)`;
`r worst_region$region` trails at `r sprintf("%+d", worst_region$nps)`. Read
the bases in the table before acting on any one region: the smaller cells carry
margins of error wide enough to overlap (see
[Methodology](#methodology-and-precision)).

## What drives the score

`ipm_model()` regresses recommendation on the rating block and crosses
importance against delivered performance. The quadrant to act on is upper
left: features that matter and are not being delivered.

```{r}
#| fig-height: 6
plot_ipm(ipm)
```

`r top_driver$feature` carries the most weight in the model at
`r round(top_driver$importance)`% of explained importance. Points are coloured
by the decision band they fall in, so anything sitting in the red zone is
below the acceptable threshold regardless of how close it looks to the next
band.

The full ranking:

```{r}
ipm[order(-ipm$importance), ]
```

Feature-level ratings show the same picture in distribution form, ordered by
weighted mean:

```{r}
rating_long <- podracing_survey %>%
  select(starts_with("ratings_")) %>%
  pivot_longer(everything(), names_to = "feature", values_to = "level") %>%
  filter(level != "") %>%
  mutate(feature = sub("^ratings_", "", feature),
         level = paste0(recode_likert(level), " - ", level)) %>%
  count(feature, level) %>%
  group_by(feature) %>%
  mutate(pct = n / sum(n) * 100) %>%
  ungroup()

plot_stacked_rating(rating_long, feature, level)
```

```{r}
#| eval: !expr params$ai
#| output: asis
#| echo: false
cat(ai_summarise(ipm, template = "drivers_barriers", context = ctx))
```

### Movement against a benchmark

With a previous wave, `compare_values()` and `plot_diff()` show the shift.
Here the mechanics are illustrated against the scale midpoint, which doubles
as a read on how far each feature sits from acceptable:

```{r}
compare_values(
  current  = ipm %>% select(feature, performance),
  previous = ipm %>% mutate(performance = threshold) %>%
    select(feature, performance)
) %>%
  plot_diff()
```

## Who the audience is

`r round(demo$pct[1])`% of respondents are
`r tolower(as.character(demo[[1]][1]))`, average age
`r round(mean_age)`.

```{r}
calc_percentage(demo_gender, sort = "desc") %>% plot_bars()
```

Occupation spreads across many long labels, so the chart switches itself to
horizontal bars:

```{r}
calc_percentage(demo_job, sort = "asc") %>% plot_bars()
```

Motivations are a check-all-that-apply block, counted over respondents rather
than ticks, so the percentages intentionally sum past 100:

```{r}
motiv %>% plot_bars(label = option)
```

`r round(motiv$pct[1])`% come for the `r motiv$option[1]`, ahead of
`r motiv$option[2]` at `r round(motiv$pct[2])`%. `r motiv$option[nrow(motiv)]`
is the minority motive, which bears on how heavily that category should front
the sponsorship mix.

A crosstab turns any pair of questions into a report-ready table:

```{r}
crosstab(region, demo_gender, cell = "row_pct")
```

## In their own words

`sample_comments_diverse()` selects maximally different verbatims. These are
illustrative of the range of opinion, not a measurement of how common each
view is.

```{r}
#| fig-height: 6
sample_comments_diverse(nps_com, show_com, n = 8, seed = 42) %>%
  plot_quotes_tree()
```

```{r}
#| eval: !expr params$ai
#| output: asis
#| echo: false
cat(ai_summarise(sample_comments_diverse(nps_com, show_com, n = 20, seed = 42),
                 template = "thematic_analysis", context = ctx))
```

## Weighting the sample

The achieved sample is `r round(demo$pct[1])`%
`r tolower(as.character(demo[[1]][1]))`. Where the fan population is known to
differ, `set_weights()` rakes the sample to the targets and every summary
helper picks the weights up automatically, reporting `wpct` beside the
unweighted `pct`:

```{r}
set_weights(c(variable = "demo_gender",
              Male = 0.49, Female = 0.50, "Non-binary" = 0.01))
calc_percentage(satis_return, sort = "desc")
clear_weights()
```

## Methodology and precision {#methodology-and-precision}

`r fmt(n_total)` responses. `precision_summary()` reports the margin of error
each headline figure carries; quote it whenever two percentages are compared,
because differences smaller than the margin are not differences.

```{r}
precision_summary(demo_gender, starts_with("ratings_"))
```

```{r}
#| eval: !expr params$ai
#| output: asis
#| echo: false
cat(ai_summarise(diagnose(demo_gender, starts_with("ratings_")),
                 template = "methodology", context = ctx))
```

```{r cleanup, include=FALSE}
clear_dataset()
```

> Rendered with `quarto render podracing-report.qmd`. Add `-P ai:true` to
> draft the AI narrative sections (needs `ellmer` and a stored key, see
> `set_llm_key()`). The matching slide deck is built by `podracing-deck.R` in
> the same folder.
