---
title: "Getting started with ezrmodel"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with ezrmodel}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 4
)
```

`ezrmodel` makes the advanced-analytics moves a research or insights analyst
reaches for into single-line helpers that return a *complete* result -- the
answer **and** its diagnostics. This tour uses the bundled datasets.

```{r setup}
library(ezrmodel)
```

## A default dataset

Set the working dataset once and call helpers without repeating it (or pipe it
in / name the argument).

```{r}
use_dataset(nps_drivers)
```

## What drives the target?

`drivers()` ranks every predictor against a target by several methods at once
(correlation, regression, relative weights, random forest, single-factor
loading) and combines them into a consensus. Methods whose package isn't
installed are simply skipped.

```{r}
d <- drivers(nps_drivers, nps)
d
```

```{r}
tidy(d)
```

```{r}
plot(d)
```

## Correlations

```{r}
correlations(nps_drivers, nps)
```

## Regression, tidily

```{r}
fit <- model_lm(nps_drivers, nps ~ quality + value + service + trust)
fit
```

```{r}
tidy(fit)
```

## Choosing and comparing models

`model_select()` lets the model pick its own predictors -- stepwise by AIC, or a
penalised lasso/ridge/elastic-net fit (via `glmnet`) that shrinks weak terms.

```{r}
step <- model_select(nps_drivers,
  nps ~ value + quality + service + ease + support + trust + price + innovation,
  method = "stepwise")
step
```

`compare_models()` lines candidates up on common metrics, best first.

```{r}
compare_models(simple = fit, stepwise = step)
```

## Clustering with diagnostics

`cluster()` returns the assignments, the sizes, the cluster profiles and the
quality diagnostics (silhouette, within-sum-of-squares) in one object.

```{r}
cl <- cluster(ecommerce, k = 4,
              vars = c(recency_days, frequency, monetary, tenure_months, returns))
cl
```

```{r}
plot(cl)
```

Read the clusters against a known variable:

```{r}
cluster_profile(cl, region)
```

## Dimensionality reduction

```{r}
r <- reduce_dims(ecommerce,
                 vars = c(recency_days, frequency, monetary, tenure_months, returns))
r
```

```{r}
plot(r)
```

For *seeing* structure rather than interpreting it, `reduce_dims()` also offers
non-linear embeddings -- `method = "umap"` (via `uwot`) or `"tsne"` (via
`Rtsne`) -- which place similar rows together in two dimensions, handy to plot
coloured by a `cluster()` assignment.

## Group tests

`test_groups()` runs a test per group -- against a benchmark or a reference
group -- and returns a tidy significance table.

```{r}
test_groups(nps_drivers, nps, region, mu = 6)
```

## Latent structure

Do the rated attributes hang together, and what dimensions sit behind them?
`reliability()` reports Cronbach's alpha; `factors()` runs an exploratory factor
analysis (choosing the number of factors for you when you don't say). Both need
the `psych` package.

```{r, eval = requireNamespace("psych", quietly = TRUE)}
reliability(nps_drivers, items = c(value, quality, service, ease,
                                   support, trust, price, innovation))
```

```{r, eval = requireNamespace("psych", quietly = TRUE)}
f <- factors(nps_drivers, vars = c(value, quality, service, ease,
                                   support, trust, price, innovation), n = 2)
f
tidy(f)
```

`sem()` confirms a structure you specify in lavaan syntax (needs `lavaan`).

## Text mining

Turn an open-text column into tidy terms, then read off the words that
distinguish groups with tf-idf. These need no extra packages.

```{r}
tokenize_text(reviews, text, review_id)
```

```{r}
term_freq(reviews, text, by = rating, top = 6)
```

For latent themes, `topics()` fits a Structural Topic Model (via `quanteda` and
`stm`); `summarise_text()` returns the most central sentences (via `lexRankr`).

```{r, include = FALSE}
clear_dataset()
```

## Where to next

* Set persistent defaults (e.g. a `seed`) with `ezrmodel_options()` and a
  profile (`use_ezrmodel_profile()`).
* Save any result table or plot with `save_data()` / `save_plot()` /
  `export_xlsx()`.
* Draft a narrative with `ai_summarise()` (needs an API key via
  `set_llm_key()`).
